\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{enumitem}
\usepackage[T1]{fontenc}

\title{COMP2300 Set 1 Practice Exam}
\author{Topics: Representation, Number Systems, Logic, CMOS Basics}
\date{}

\begin{document}
\maketitle

\noindent\textbf{Time:} 120 minutes \hfill
\textbf{Total:} 100 marks

\vspace{0.5em}
\noindent\textbf{Instructions}
\begin{itemize}[leftmargin=1.5em]
  \item Answer in English.
  \item Part A is multiple-choice (concept-focused).
  \item Part B is computational. Show all working for full credit.
  \item Assume fixed-width arithmetic where width is explicitly given.
\end{itemize}

\section*{Part A: Multiple Choice (30 marks, 3 marks each)}
Choose exactly one option for each question.

\begin{enumerate}[label=\textbf{A\arabic*.}]
  \item In the transformation hierarchy, which is the hardware/software interface?
  \begin{enumerate}[label=(\Alph*)]
    \item Algorithm
    \item ISA
    \item Microarchitecture
    \item Circuits
  \end{enumerate}

  \item Which statement about ISA vs. microarchitecture is correct?
  \begin{enumerate}[label=(\Alph*)]
    \item ISA is how transistors are wired.
    \item Microarchitecture is the programmer-visible instruction definition.
    \item Different microarchitectures can implement the same ISA.
    \item ISA changes every time clock frequency changes.
  \end{enumerate}

  \item Why is binary preferred in digital systems?
  \begin{enumerate}[label=(\Alph*)]
    \item It always uses less memory than decimal.
    \item It simplifies hardware and improves reliability.
    \item It removes the need for abstraction.
    \item It avoids all overflow.
  \end{enumerate}

  \item Which is true for 2's complement representation?
  \begin{enumerate}[label=(\Alph*)]
    \item It has two representations for zero.
    \item It cannot represent negative numbers.
    \item Ordinary binary addition can be used.
    \item Overflow is detected only by carry-out.
  \end{enumerate}

  \item For 2's complement addition, overflow occurs when:
  \begin{enumerate}[label=(\Alph*)]
    \item Carry-out from MSB is 1.
    \item The addends have different signs.
    \item The addends have the same sign and result has opposite sign.
    \item Result is zero.
  \end{enumerate}

  \item Sign extension of an $N$-bit 2's complement value to $M$ bits ($M>N$):
  \begin{enumerate}[label=(\Alph*)]
    \item Always appends zeros.
    \item Always appends ones.
    \item Replicates the sign bit into new high bits.
    \item Inverts all bits.
  \end{enumerate}

  \item In CMOS logic, a valid steady state ideally has:
  \begin{enumerate}[label=(\Alph*)]
    \item Pull-up ON and pull-down ON.
    \item Pull-up OFF and pull-down OFF.
    \item Exactly one of pull-up/pull-down ON.
    \item Both ON if output is 1.
  \end{enumerate}

  \item Which function outputs 1 when inputs are different?
  \begin{enumerate}[label=(\Alph*)]
    \item AND
    \item OR
    \item XOR
    \item XNOR
  \end{enumerate}

  \item A bit mask is primarily used to:
  \begin{enumerate}[label=(\Alph*)]
    \item Increase clock frequency.
    \item Select/clear specific bits.
    \item Convert analog to digital.
    \item Remove sign extension.
  \end{enumerate}

  \item A half adder computes:
  \begin{enumerate}[label=(\Alph*)]
    \item Sum of two bits with carry-in.
    \item Sum and carry for two input bits (no carry-in).
    \item Product and quotient.
    \item Bitwise NOR and NAND.
  \end{enumerate}
\end{enumerate}

\section*{Part B: Computational Problems (70 marks)}
\begin{enumerate}[label=\textbf{B\arabic*.}]
  \item \textbf{Information content and encoding (5 marks)}\\
  (a) How many bits are minimally required to encode 26 symbols?\\
  (b) How many bits are minimally required to encode 100 symbols?

  \item \textbf{Base conversion I (6 marks)}\\
  Convert decimal \(53_{10}\) to binary using any valid method.

  \item \textbf{Base conversion II (6 marks)}\\
  Convert binary \(11010110_2\) to decimal.

  \item \textbf{Binary/hex conversion (6 marks)}\\
  (a) Convert \(1111\,0111_2\) to hexadecimal.\\
  (b) Convert \(D741_{16}\) to binary.

  \item \textbf{Representable ranges (6 marks)}\\
  For \(N=8\), give the min and max for:\\
  (a) Unsigned\\
  (b) Sign/magnitude\\
  (c) 2's complement

  \item \textbf{2's complement encoding/decoding (8 marks)}\\
  Using 8 bits:\\
  (a) Encode \(-18\) in 2's complement.\\
  (b) Decode \(1110\,1011_2\) as a signed 2's complement integer.

  \item \textbf{Unsigned addition and overflow (6 marks)}\\
  Add \(1111_2 + 1111_2\) as 4-bit unsigned values.\\
  Provide 4-bit result and state whether overflow occurs.

  \item \textbf{2's complement addition and overflow (8 marks)}\\
  Use 5-bit 2's complement arithmetic.\\
  (a) \(01001_2 + 01011_2\)\\
  (b) \(10100_2 + 11010_2\)\\
  For each, report binary result, decimal interpretation, and whether overflow occurs.

  \item \textbf{Sign extension (5 marks)}\\
  Sign-extend \(1101_2\) (4-bit 2's complement) to 8 bits, then give its decimal value.

  \item \textbf{Bitwise operations (6 marks)}\\
  Let \(A=00110110_2\), \(B=00001111_2\). Compute:\\
  (a) \(A\ \&\ B\)\\
  (b) \(A\ |\ B\)\\
  (c) \(A\ \oplus\ B\)

  \item \textbf{Bit mask tasks (6 marks)}\\
  Let \(X=10110010_2\).\\
  (a) Extract the lowest 4 bits of \(X\) (show mask and result).\\
  (b) Clear the lowest 2 bits of \(X\) (show mask and result).\\
  (c) Set the highest 2 bits of \(X\) to 1 (show mask and result).

  \item \textbf{Logic/truth table and useful circuits (8 marks)}\\
  (a) For \(Y=(A+B)'\), complete the 2-input truth table (A,B,Y).\\
  (b) A 2:1 MUX has inputs \(I_0=1\), \(I_1=0\), select \(S\). Write \(Y\) in terms of \(S\), then evaluate for \(S=0\) and \(S=1\).\\
  (c) For a half adder with inputs \(A=1, B=1\), compute Sum and Carry.
\end{enumerate}


\section*{Part C: Past Exam Questions}

\begin{enumerate}[label=\textbf{C\arabic*.}]
  \item \textbf{Digital Logic Fundamentals (from 2023 Final Exam)}
  \begin{itemize}
    \item During a work assignment, you are asked to design a combinational circuit with a three-bit input, \{A, B, C\} (A is the most significant bit and C is the least significant bit), and two 1-bit outputs, Factorial and Div3. The value of each output is determined as follows:
    \begin{itemize}
        \item The output Factorial is 1 if the input number is equal to its factorial. Factorial is the product of all positive integers (excluding 0) that are less than or equal to the input number. When the input combinations are all 0, the output is 1.
        \item The output Div3 is 1 only when the input 3-bit number is divisible by 3.
    \end{itemize}
    \item Answer the following questions.
    \begin{enumerate}
        \item[A.] Draw the truth table for the combinational circuit.
        \item[B.] Express the output Factorial as the simplest sum of products representation. Show your work step-by-step.
        \item[C.] Express the output Div3 as the simplest sum of products representation. Show your work step-by-step.
        \item[D.] Draw a schematic of Factorial and Div3 using basic logic gates.
        \item[E.] It is well-known that we can implement any Boolean function either using basic logic gates or a look-up table (N to 1 multiplexer). Explain the tradeoff between the two approaches.
    \end{enumerate}
  \end{itemize}

  \item \textbf{Digital Logic Fundamentals (from 2023 Final Exam Supp/DA)}
  \begin{itemize}
    \item During a work assignment, you are asked to design a combinational circuit with a three-bit input, \{A, B, C\} (A is the most significant bit and C is the least significant bit), and two 1-bit outputs, O and E. The value of each output is determined as follows:
    \begin{itemize}
        \item The output E is 1 if the input number is even. When the input combinations are all 0, the output is 1.
        \item The output O is 1 only when the input 3-bit number is odd.
    \end{itemize}
    \item Answer the following questions.
    \begin{enumerate}
        \item[A.] Draw the truth table for the combinational circuit.
        \item[B.] Express the output E as the simplest sum of products representation. Show your work step-by-step.
        \item[C.] Express the output O as the simplest sum of products representation. Show your work step-by-step.
        \item[D.] Draw a schematic of E and O using basic logic gates.
        \item[E.] Explain the difference between a multiplexer and a decoder. Which one of the two has a higher logic complexity?
    \end{enumerate}
  \end{itemize}

  \item \textbf{Multiple Choice (from exam24px1.pdf)}
    \begin{enumerate}
        \item How many bits are needed to represent -512 with sign/magnitude numbers?
        \item If one of the inputs of an XNOR gate is connected to logic level 1 (TRUE) forever, the resulting circuit behaves like: (OR gate, NOR gate, NOT gate, Short circuit wire, None of the above)
        \item Under De Morgan's Theorem, which of the following circuits are equivalent: (An NOR gate with inverted output is equivalent to an AND gate with one inverted input, An AND gate with inverted output is equivalent to a NOR gate with inverted output, An AND gate with inverted output is equivalent to an OR gate with inverted inputs, An OR gate with inverted output is equivalent to an OR gate with inverted inputs, A NAND gate with inverted output is equivalent to an OR gate with inverted inputs, None of the above)
        \item A 2:4 decoder can be used to implement an XNOR gate by connecting which ones of its one-hot outputs to the inputs of an OR gate: (00 and 01, 00 and 11, 01 and 10, 10 and 11, 00 and 10)
    \end{enumerate}

\end{enumerate}

\end{document}

